% =================
% Preface page
% =================
\chapter{Preambule}

\lettrine{I}{have spent years building} devices based on Raspberry Pi: time attendance machines, camera boards, thermal vision systems, access control devices, even a self-service beer tap (Cybertap).


For a long time, I wished there was a book like this one. A book that didn’t just show the theory, but told the truth about what it really takes to move from an idea to a production-ready product.

I made plenty of mistakes along the way. I chose the wrong boards, I burned Raspberry Pis, I lost months fixing devices. At one point, I was spending eight hours a day repairing customer units because my devices would only last one year — while I was offering a two-year warranty. The biggest lesson? Never rely on SD cards in production.

Despite the failures, I kept going. Over time, I sold more than 5000 of my devices worldwide. Every project taught me something new, and those lessons are what I want to share with you here.

If you want to create devices, put them into the market, and avoid the mistakes I made — this book is for you.

\section*{About Author}

I’m Adam, and I love to create, program, and build. I’ve always been curious about technology and people, and I enjoy not only designing but also selling what I create.

My first company was born while I was still at university—StudentsCode. Back then, I ran Minecraft servers that surprisingly brought me both a good income and my future wife. Later, I worked at Siemens as a Software Architect, and then at the Texas startup Codenotary, where I grew from Software Developer to Chief Architect over the course of three years.

Eventually, I returned to what had always been close to me: Raspberry Pi. I began producing my own devices, learning the hard way what it means to build for real customers and ship hardware worldwide.

Today, I have designed and delivered thousands of devices based on Raspberry Pi technology, ranging from time attendance systems to access control units, thermal vision cameras, and self-service beer machines. This book collects the lessons I’ve learned along that journey.

\cleardoublepage   % Make sure contents page starts on right-side page